HTW E1 Testing

### 2023-09-30 Analysis/e1_test.qmd
render changes
re-render

To assess differences between groups, we used Bayesian Mixed Effects
Regression. Model fitting was performed with the brms package in R
@burknerBrmsPackageBayesian2017, and descriptive stats and tables were
extracted with the BayestestR package
@makowskiBayestestRDescribingEffects2019a. Mixed effects regression
enables us to take advantage of partial pooling, simultaneously
estimating parameters at the individual and group level. Our use of
Bayesian, rather than frequentist methods allows us to directly quantify
the uncertainty in our parameter estimates, as well as circumventing
convergence issues common to the frequentist analogues of our mixed
models. For each model, we report the median values of the posterior
distribution, and 95% credible intervals.

@@

\begin{equation}
vx_{ij} = \beta_0 + \beta_1 \cdot condit_{ij} + \beta_2 \cdot bandInt_{ij} + \beta_3 \cdot condit_{ij} \cdot bandInt_{ij} + b_{0i} + b_{1i} \cdot bandInt_{ij} + \epsilon_{ij}
\end{equation}

```{r}
#| label: fig-e1-test-vx
#| fig-cap: E1 testing x velocities. Translucent bands with dash lines indicate the correct range for each velocity band. 
#| fig-width: 11
#| fig-height: 9
test %>% group_by(id,vb,condit) |> plot_distByCondit()

@@

See @tbl-e1-bmm-vx for the full model results. The estimated coefficient
for training condition (β = 164.67, 95% CrI [45.50, 278.85]) suggests
that the varied group tends to produce harder throws than the constant
group, but is not in and of itself useful for assessing discrimination.
Most relevant to the issue of discrimination is the slope on Velocity
Band (β = 0.71, 95% CrI [.62, .80]). Although the median slope does fall
underneath the ideal of value of 1, the fact that the 95% credible
interval does not contain 0 provides strong evidence that participants
exhibited some discrimination between bands. The slope coefficient (β =
0.71 ) also seems in accordance with an examination of
@tbl-e1-test-nf-vx, which shows that average x-velocity fell within the
range of the target band for four out of six of the bands ($4/6$=.67).
The estimate for the interaction between slope and condition (β = -.12,
95% CI [-.25, 0.00]), suggests that the varied condition had
significantly lower slopes than did the constant group. This difference
is depicted visually in @fig-e1-bmm-vx.

@@

indv_coefs <- coef(e1_vxBMM)$id |> 
  as_tibble(rownames="id") |> 
  select(id, starts_with("Est")) |>
  left_join(e1Sbjs, by=join_by(id) )

@@

  indvSlopes |> group_by(condit) |> reframe(enframe(quantile(Slope, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "Slope")) |> 
  pivot_wider(names_from=quantile,values_from=Slope,names_prefix="Q_") |>
  group_by(condit) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kbl()

@@

indvDraws |> group_by(condit) |> median_qi(SlopeInt)

e1_vxBMM |> emmeans(~condit) |> gather_emmeans_draws()

 ggplot(indvDraws, aes(x = SlopeInt, color = condit)) + 
  geom_density()

@@

```{r}
#| label: fig-e1-bmm-bx2
#| fig-cap: Slope estimates by participant - ordered from lowest to highest within each condition. 
#| fig-height: 9
#| fig-width: 10

@@

```{r}
#| label: Band Deviation
#| tbl-cap: "Deviations from target"
#| layout: [[70, 30],[-5]]

>>>>

All data processing and statistical analyses were performed in R version 4.31 @rcoreteamLanguageEnvironmentStatistical2020. To assess differences between groups, we used Bayesian Mixed Effects Regression. Model fitting was performed with the brms package in R @burknerBrmsPackageBayesian2017, and descriptive stats and tables were extracted with the BayestestR package @makowskiBayestestRDescribingEffects2019a. Mixed effects regression enables us to take advantage of partial pooling, simultaneously estimating parameters at the individual and group level. Our use of Bayesian, rather than frequentist methods allows us to directly quantify the uncertainty in our parameter estimates, as well as circumventing convergence issues common to the frequentist analogues of our mixed models. For each model, we report the median values of the posterior distribution, and 95% credible intervals.

@@

```{=tex}
\begin{equation}
vx_{ij} = \beta_0 + \beta_1 \cdot condit_{ij} + \beta_2 \cdot bandInt_{ij} + \beta_3 \cdot condit_{ij} \cdot bandInt_{ij} + b_{0i} + b_{1i} \cdot bandInt_{ij} + \epsilon_{ij}
\end{equation}
```
```{r}
#| label: fig-e1-test-vx
#| fig-cap: E1 testing x velocities. Translucent bands with dash lines indicate the correct range for each velocity band. 
#| fig-width: 11
#| fig-height: 9
test %>% group_by(id,vb,condit) |> plot_distByCondit()

@@

See @tbl-e1-bmm-vx for the full model results. The estimated coefficient for training condition (β = 164.67, 95% CrI [45.50, 278.85]) suggests that the varied group tends to produce harder throws than the constant group, but is not in and of itself useful for assessing discrimination. Most relevant to the issue of discrimination is the slope on Velocity Band (β = 0.71, 95% CrI [.62, .80]). Although the median slope does fall underneath the ideal of value of 1, the fact that the 95% credible interval does not contain 0 provides strong evidence that participants exhibited some discrimination between bands. The slope coefficient (β = 0.71 ) also seems in accordance with an examination of @tbl-e1-test-nf-vx, which shows that average x-velocity fell within the range of the target band for four out of six of the bands ($4/6$=.67). The estimate for the interaction between slope and condition (β = -.14, 95% CI [-.26, -0.01]), suggests that the discrimination was somewhat modulated by training condition, with the varied participants showing less senitivity between vands than the constant condition. This difference is depicted visually in @fig-e1-bmm-vx.

The second model, which focused solely on extrapolation bands, revealed similar patterns. The  Velocity Band term (β = 0.50, 95% CrI 0.42,0.570.42,0.57) still demonstrates a high degree of discrimination ability. However, the posterior distribution for interaction term (β = -0.07, 95% [CrI −0.17,0.04] ) does across over 0, suggesting that the evidence for decreased discrimination ability for the varied participants is not as strong when considering only the three extrapolation bands.

@@

indv_coefs <- as_tibble(coef(e1_vxBMM)$id, rownames="id")|> 
  select(id, starts_with("Est")) |>
  left_join(e1Sbjs, by=join_by(id) )

@@

@@

```{r}
#| label: fig-e1-bmm-bx2
#| fig-cap: Slope distributions between condition
#| fig-subcap: ["Slope estimates by participant - ordered from lowest to highest within each condition. ", "Model fit to only 3 extrapolation bands"]
#| fig-height: 9
#| fig-width: 10

@@

@@
